\textbf{Modelling Neuroplasticity: Using Machine Learning to Learn About TMS Data}

### 2023-06-11 paper/paper.tex
.

The distribution of the evoked potentials was always exponential. However, the
$\beta$ parameter of the distribution varies across subject groups and session
types, as shown in Figures \ref{fig:figure1} and \ref{fig:figure2} and in the
table below:

@@

For simplicity, we deal with the hypothetical situation in which a single ISI is
used for paired stimulations in a unique TMS session. Our formulations
generalize easily to different inter-stimulus intervals and multiple
experimental sessions.

Firstly, we need to establish some convenient notation. Let $n$ and $m$ be the
number of paired and test pulses elicited to the subject in the session,
respectively. Let $\textbf{x}$, $\textbf{t}$ be vectors over the $\mathbb{R}^+$
vector space whose components are the evoked potentials of paired and pulse
stimulations, respectively.

@@

    Let $x \in \textbf{x}$ be the evoked potential of a single paired
    stimulation. Then we define two pulse-specific relative amplitude measures,

@@

@@

On the contrary, the inclusion of weighted variants of the features defined
above greatly improves the performance of the model. The use of weighted instead
of arithmetic averages in the formulas for $\delta$ and $\rho$ is useful in
dealing with the excessive influence of outliers or highly spread out points in
the feature. For example, the weight vectors may be computed using the MAD or
the inverse-variance of each $x$.

@@

>>>>

The distribution of the evoked potentials is always exponential. However, the
$\beta$ parameter of the distribution varies across subject groups and session
types, as shown in Figures \ref{fig:figure1} and \ref{fig:figure2} and in the
table below:

@@

For simplicity, we will first deal with the hypothetical situation in which a
single ISI is used for paired stimulations in a unique TMS session. Our
formulations generalize easily to different inter-stimulus intervals and
multiple experimental sessions.

Firstly, we need to establish some convenient notation. Let $n$ and $m$ be the
number of paired and test pulses elicited to the subject in the session,
respectively. Let $\textbf{x}$, $\textbf{t}$ be vectors over the $\mathbb{R}^+$
vector space whose components are understood to be the evoked potentials of
paired and test stimulations, respectively.

@@

    Let $x \in \textbf{x}$ be the evoked potential of a single paired
    stimulation. Then we define two pulse-specific relative amplitude measures,
    $\rho : \mathbb{R} \to \mathbb{R}, \delta : \mathbb{R} \to \mathbb{R}$,
    where

@@

\begin{remark}
    Let $\pi$ denote the coefficient traditionally used for estimating the
    relative amplitude between test and paired pulses. In accordance to what was
    described in the second paragraph of Section $4$,

    \begin{align*}
        \pi = \frac{\mu_x}{\mu_t}
    .\end{align*}

    where $\mu_x, \mu_t$ are the mean paired and test pulse responses,
    respectively. It follows from the definition of $\pi$ and $\rho$ that

    \begin{align*}
        \pi = \frac{\rho(x_1) + \ldots + \rho(x_n)}{n}
    .\end{align*}

    (For a proof, consult the appendix.)
\end{remark}

The second remark is of great importance. It states that the way in which
relative amplitude was traditionally computed---the way in which neuroplasticity
was traditionally proxied---is different from our proposed $\rho$ function only
in terms of scale. More specifically, the traditional coefficient used to proxy
neuroplasticity at the subject level is nothing but the mean $\rho$
measure in an experimental session. We believe this bestows $\rho$ with a
certain degree of confidence, insofar as not only it is not entirely new, but
neuroscientists have implicitly used it all along.

\subsection{Generalizing to many subjects}

\begin{definition}
    Let $\textbf{x} \in \mathbb{R}^j$. Then we define $T(\textbf{x}) := \begin{bmatrix} 
        \rho(x_1) & \ldots & \rho(x_j)
    \end{bmatrix}^\intercal $ and $R(\textbf{x}) := \begin{bmatrix} 
        \delta(x_1) & \ldots & \delta(x_j) 
    \end{bmatrix}^\intercal $.
\end{definition}

\begin{theorem}
    The mappings $T, R$ are linear transformations.
\end{theorem}

Let $k \in \mathbb{N}$ denote the number of experimental subjects in an
experiment, to each of whom $n, m$ paired and test pulses were elicited,
respectively.

\begin{definition}
    We define $\textbf{P} \in \mathbb{R}^{k \times n}$ a
    matrix where the $i$th row vector $\textbf{x}^{(i)}$ is the vector whose
    components are the paired-pulse responses of the $i$th subject.
\end{definition}

\begin{definition}
    Let $\mu^{(t)}_i$ be average test-pulse response of the $i$th subject. Then
    we define
    $\textbf{M}^{k \times k}$ a diagonal matrix with $\textbf{M}_{(ii)} =
    \frac{1}{\mu^{(t)}_{i}}$.
\end{definition}

\begin{theorem}
    The row-vectors of the matrix $\textbf{X}_\rho = \textbf{M}\textbf{P}$ are the
    vectors $T(\textbf{x}^{(1)}), \ldots, T(\textbf{x}^{(k)})$.
\end{theorem}

Similarly, if we let $\textbf{M'}$ be a diagonal matrix such that
$\textbf{M'}_{ii} = \mu^{(t^{-1})}_i$, where $\mu^{(t^{-1})}_i$ denotes the
average \textit{inverse} test-pulse response of the $i$th subject, we come to
the following result.

\begin{theorem}
    The row-vectors of the matrix $\textbf{X}_\delta = \textbf{M'}\textbf{P}$ 
    are the vectors $R(\textbf{x}^{(1)}), \ldots,
    R(\textbf{x}^{(k)})$.
\end{theorem}

If experiments were conducted using multiple ISI, then to each ISI $r$ it
corresponds a distinct matrix $\textbf{X}_\delta^{ISI = r},
\textbf{X}_\rho^{ISI = r}$. The same applies to different subject groups.

We have expanded $\rho, \delta$ to higher dimensions for the purpose of
providing a theoretical framework of their application to larger data sets and
many experimental subjects. However, it is important to note that the matrix
representations here presented only make sense under the assumption that, on
each subject, none of $n$ paired pulses were anomalous, and hence all can be
safely included in the model. Such assumption rarely holds in reality, where it
is usual to excluded some pulses due to the presence of artifacts. There are
many ways to deal with this situation in the implementation of our model---the
safest being the direct application of $T$ and $R$ to the data vectors---. Thus,
more than serving the algorithmic implementation of the model, which may
prescind from them, these matrices provide insight into the nature of the
transformations involved in the model.

@@

On the contrary, the inclusion of weighted variants of the features defined
above greatly improves the performance of the model. The use of weighted instead
of arithmetic averages in the formulas for $\delta$ and $\rho$ is useful in
dealing with the excessive influence of outliers or highly spread out points in
the feature. For example, the weight vectors may be computed using the MAD or
the inverse-variance of each $t \in \textbf{t}$.

@@

\textbf{Proof 2.} In \textbf{Remark 2}, we observe that $\rho$ is implicitly
used in the traditional measures of relative amplitude. Recall that such methods
consist in taking the average paired-pulse response and dividing it by the
average test-pulse response. It is easily observed that

\begin{align*} \frac{\mu_x}{\mu_d} = \frac{\frac{1}{n}\sum_{j=1}^n
        x_{j}}{\frac{1}{m}\sum_{j = 1}^m t_j} 
        = \frac{\rho(x_{1i})+\cdots + \rho(x_{ni})}{n}
\end{align*}

\begin{proof}[Theorem 1]

    Let $\sigma = \sum_{i=1}^m t_j$ and $c \in \mathbb{R}$. It follows directly
    from the definition of $\rho$ that

    \begin{align*}
        T(c\textbf{x} + \textbf{y}) &= \begin{bmatrix} \frac{m(cx_1) + y_1}{\sigma} &
        \ldots& \frac{m(cx_n + y_n)}{\sigma} \end{bmatrix}^\intercal \\ 
                                    &=\begin{bmatrix} 
                                        \frac{mcx_1}{\sigma} +
                                        \frac{my_1}{\sigma} & \ldots &
                                        \frac{mcx_n}{\sigma} + \frac{y_n}{\sigma} 
                                    \end{bmatrix} \\
                                    &=cT(\textbf{x}) + T(\textbf{y})
    \end{align*}

    The proof the $R$ is linear is practically identical.
\end{proof}
